### 2008-02-19 SAiP/Module1/H2/table.tex
Adding LaTex document with the first scenario begun

\documentclass[10pt,a4paper]{article}

\usepackage{times}
\usepackage[latin1]{inputenc}
\usepackage{rotating}
\usepackage{multirow}

\begin{document}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & After an average user has used the system for less than one hour e or she has mastered the basic functionality \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} usability & \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} &  \\
  \cline{2-3}
  & {\bfseries Stimulus:} &  \\
  \cline{2-3}
  & {\bfseries Artifact} &  \\
  \cline{2-3}
  & {\bfseries Environment:} &  \\
  \cline{2-3}
  & {\bfseries Response:} &  \\
  \cline{2-3}
  & {\bfseries Response Measure:} & \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} &  \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Scenario refinement for HS07-...}
\end{center}
\end{table}

\end{document}