### 2008-02-28 trunk/SAiP/Module1/H3/doc/beerwebstore.tex
Each section moved into seperate .tex file.

%%%%
%%%% Beer Web Store
%%%%
\section{Software Architecture in Practice Case: Beer Web Store}