### 2008-02-28 SAiP/Module1/H3/doc/architecturaldescription.tex
Each section moved into seperate .tex file.

%%%
% Architectural Description
%%%
\subsection{Architectural Description -- Beer Web Store}
The architect for a new Beer Web Store for specialty beers has drawn
the following figure of the architecture together with his customers.
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}
The system is to support web users that may purchase beers using a web
browser. The system runs on top of a database (``DB''), contains
domain model (``Data''), an overview page (``Overview''), and a
``Basket'' subpart.

Discuss:
\begin{itemize}
    \item Which structures does a system such as the above exhibit?
    Give examples of elements and relations pertaining to each
    structure
    \item Which viewpoints are relevant when describing the example?
    Give examples of partials views for each viewpoint
    \item Argue for benefits and liabilities of describing software
    architecture via a box-and-line drawing such as the above
    \item Discuss what architectural description would be needed if
    the system was to be realized as a Service-Oriented Architecture
\end{itemize}