### 2008-02-28 SAiP/Module1/H3/doc/softwarearchitecture.tex

%%%
% Architectural Description
%%%
\subsection{Software Architecture -- Beer Web Store}
A new Beer Web Store for specialty beers has the following main use
case:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated and he is presented with a receipt
\end{quote}
As a first step, the software architect of this systems has decided to
use a layered architectural style (with clients, server, and a
database in separate layers). He has decided that the communication to
the database will be done via a Fa\c cade pattern. The system will be
built in Java.

Discuss:
\begin{itemize}
    \item Outline how you would approach the tasks of creating the
    architecture for the Beer Web Store. Consider, e.g., which steps
    would you takes and in which order?
    \item Relate the definition of 'software architecture' in [Bass et
    al., 2003] to the elements of your design as outlined above. Give
    concrete examples of elements, relations, and structures.
    \item Apply [Perry and Wolf, 1992]'s model of software architecture as
      \begin{quote}
	{\it Software Architecture = \{Elements, Form, Rationale\}}
      \end{quote}
      to the Beer Web System
    \item Reflect on what happens if the words ``software'' and
    ``computing'' are removed from the definition of 'software
    architecture' in [Bass et al., 2003]
    \item The architect decides to create a full architecture
    description before embarking on any implementation of the
    system. Discuss pros and cons of taking that approach
\end{itemize}