### 2008-03-01 SAiP/Module1/H3/doc/architecturaldescription.tex
Formatting of questions completed.

%%%
% Architectural Description
%%%
\section{Architectural Description -- Beer Web Store}

The architect for a new Beer Web Store for specialty beers has drawn
the following figure of the architecture together with his customers.
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}
The system is to support web users that may purchase beers using a web
browser. The system runs on top of a database (``DB''), contains
domain model (``Data''), an overview page (``Overview''), and a
``Basket'' subpart.

\subsection{Architectural Description - Question 1}

\begin{question}
Which structures does a system such as the above exhibit?
Give examples of elements and relations pertaining to each
structure
\end{question}

\fixme{Answer question 2.1}

\subsection{Architectural Description - Question 2}

\begin{question}
Which viewpoints are relevant when describing the example?
Give examples of partials views for each viewpoint
\end{question}

\fixme{Answer question 2.2}

\subsection{Architectural Description - Question 3}

\begin{question}
Argue for benefits and liabilities of describing software
architecture via a box-and-line drawing such as the above
\end{question}

\fixme{Answer question 2.3}

\subsection{Architectural Description - Question 4}

\begin{question}
Discuss what architectural description would be needed if
the system was to be realized as a Service-Oriented Architecture
\end{question}

\fixme{Answer question 2.4}