### 2008-03-01 SAiP/Module1/H3/doc/productlines.tex

%%%
% Product Lines and Frameworks
%%%
\section{Product Lines -- Beer Web Store}
Consider a Beer Web Store for buying specialty beers. The software
architect has designed a deployment structure of the system as shown
below:
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/bws-deployment,scale=0.80}}
  \label{fig:observation}
\end{figure}
The main architectural drivers for the system are performance and
modifiability and the following is the main use case for the system:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}

The company behind the Beer Web Store has decided to build a product
line based on the store so that, e.g., sodas or stamps may be sold in
a new shop or other delivery platforms such as mobile phones may be
used.

\subsection{Product Lines - Question 1}

\begin{question}
How will the choice of a product line approach affect the
architecture of the beer web store?
\end{question}

\fixme{Answer question 5.1}

\subsection{Product Lines - Question 2}

\begin{question}
Discuss the benefits and liabilities of using a product line
approach in relation to the case?
\end{question}

\fixme{Answer question 5.2}

\subsection{Product Lines - Question 3}

\begin{question}
Relate Service-Oriented Architecture and Product Lines?
\end{question}

\fixme{Answer question 5.3}