### 2008-03-01 SAiP/Module1/H3/doc/qualityattributes.tex

%%%
% Quality Attributes
%%%
\section{Quality Attributes -- Beer Web Store}

Consider the following main use case for a Beer Web Store for specialty
beers:

\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. Moreover, the software architect for the system has
created the following box-and-line drawing documenting his initial
ideas of the system:
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}

\subsection{Quality Attributes - Question 1}

\begin{question}
Describe technique(s) for architectural requirements capture
that are applicable to the above case
\end{question}

\fixme{Answer question 3.1}

\subsection{Quality Attributes - Question 2}

\begin{question}
Give feasible architectural requirements for availability
and performance for the Beer Web Store using such techniques (at
least one for each quality attribute)
\end{question}

\fixme{Answer question 2.2}

\subsection{Quality Attributes - Question 3}

\begin{question}
Argue how tactics and/or styles may be used to resolve the requirements
\end{question}

\fixme{Answer question 3.3}

\subsection{Quality Attributes - Question 4}

\begin{question}
Discuss where Service-Oriented Architecture can play a role
with respect to the Beer Web Store? Consider the quality implications
\end{question}

\fixme{Answer question 3.4}

\subsection{Quality Attributes - Question 5}

\begin{question}
Reflect upon what the role of quality attributes in software
architecture is
\end{question}

\fixme{Answer question 3.5}