### 2008-03-17 trunk/SAiP/Module1/H3/doc/beerwebstore.tex
Jesper Dalberg removed from front page.

%%%%
%%%% Beer Web Store
%%%%

\title{Exercise 3: Software Architecture in Practice Case: Beer Web Store}

\author{
  Anders H Poder, Lars Kringelbach\\\\
  Department of Computer Science, University of Aarhus\\
  Aabogade 34, 8200 {\AA}rhus N, Denmark\\\\
  \makeatletter
  \texttt{Group 11 - Kilo}\\
  \texttt{19951439, 20074842}\\
  \texttt{\{ahp, u074842\}@daimi.au.dk}
}

\date{2008-03-17}

\maketitle